---
title: "Final Project R"
author: "Yelizaveta Nazarova"
date: "2026-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, library}
# install.packages("tidyverse")
# install.packages("stargazer")
library(tidyverse)
library(stargazer)
library(dplyr)

data <- read.csv("/Users/liza/Documents/GitHub/final project eco 590/final_dataset.csv")
```
<h3>1st Regression </h3>
```{r}
ols1 = lm(close ~ temperature + humidity + cloud_cover + wind_speed, data = data) 
summary(ols1)
# OLS1 examines the relationship between weather variables and S&P 500 closing prices.
# The purpose of this regression is to test whether weather conditions have a stronger relationship with stock market prices during unusually cold or hot periods.
```
<h3>2st Regression (with filter)</h3>
```{r}
extreme_data = subset(data, temperature < 0 | temperature > 25)

ols2 = lm(close ~ temperature + humidity + cloud_cover + wind_speed, data = extreme_data)

summary(ols2)
# OLS2 uses a filtered dataset that only includes extreme temperature days (below 0°C or above 25°C)
```
<h3>3st Regression (Preffered)</h3>
Refference: https://stats.stackexchange.com/questions/332458/how-do-you-incorporate-seasonal-differencing-into-a-regression-in-r
```{r}
data$month = format(as.Date(data$date), "%m")

ols3 = lm(close ~ bad_weather + factor(month), data = data) 
summary(ols3)
```

```{r, results='asis'}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
stargazer(
  ols1,
  ols2,
  ols3,
  type = "html",
  title = "Regression Results",
  dep.var.caption = "",   # removes dependent variables
  column.labels = c("OLS 1", "OLS 2", "OLS 3"),
  covariate.labels = c(
    "Temperature",
    "Humidity",
    "Cloud Cover",
    "Wind Speed",
    "Bad Weather"
  ),
  out = "name.html"
)
```


<h3>Interpretation of the Regression</h3>
<strong>OLS3</strong> is my preferred regression because it controls for seasonality by including monthly fixed effects using factor(month). This allows the model to separate the effect of bad weather from general monthly trends in the stock market. The regression examines the relationship between bad weather conditions and S&P 500 closing prices while holding seasonal effects constant.

The coefficient for bad_weather is -1.379. This means that, on average, days classified as bad weather are associated with approximately a 1.38-point decrease in S&P 500 closing prices, holding seasonal effects constant. However, the p-value for this variable is 0.635, which is greater than 0.05. Therefore, the relationship is not statistically significant, meaning there is not enough evidence to conclude that bad weather alone has a strong independent effect on market prices after controlling for seasonality.

Most monthly variables are statistically significant, especially from March through November. This indicates that seasonality plays an important role in explaining stock market performance. Also, the model has a high R-squared value of 0.8165, meaning that approximately 82% of the variation in S&P 500 closing prices is explained by the regression model.

